---
title: "Birthdays and Holidays"
author: "Oliver Thistlethwaite"
output: html_document
---   

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this project is to analyze any possible relationship between birthdays and holidays. We'll use the R libraries *dplyr*, *ggplot2*, *lubridate*, as well as the *Birthdays* data set from the *mosaicData* library.

```{r,include=FALSE}
library(dplyr)
library(ggplot2)
library(lubridate)
data(Birthdays, package = "mosaicData")
head(Birthdays)
```

First we will draw a plot measuring the number of births for each day in our time frame.

```{r}
DailyBirths <- Birthdays %>%
  group_by(date) %>%
  summarise(total = sum(births))

DailyBirths %>%
  ggplot(aes(x=date, y = total)) + geom_point()

```

Next we draw a similar plot, only this time measured by week of the year.

```{r}
WeeklyBirths <- Birthdays %>%
  mutate(week = week(date)) %>%
  group_by(week) %>%
  summarise(total = sum(births))

WeeklyBirths %>%
  ggplot(aes(x=week, y = total)) + geom_point()

```

Now by month of the year.

```{r}
MonthlyBirths <- Birthdays %>%
  mutate(month = month(date)) %>%
  group_by(month) %>%
  summarise(total = sum(births))

MonthlyBirths %>%
  ggplot(aes(x=month, y = total)) + geom_point()

```

Here we plot by day of the year.

```{r}
YDailyBirths <- Birthdays %>%
  mutate(yday = yday(date)) %>%
  group_by(yday) %>%
  summarise(total = sum(births))

head(YDailyBirths)

YDailyBirths %>%
  ggplot(aes(x=yday, y = total)) + geom_point()

```

Next by day of the week.

```{r}
WDailyBirths <- Birthdays %>%
  mutate(wday = wday(date, label = TRUE)) %>%
  group_by(wday) %>%
  summarise(total = sum(births))

head(WDailyBirths)

WDailyBirths %>%
  ggplot(aes(x=wday, y = total)) + geom_point()

```

For the remainder of this document, we restrict to the year 1980. First we graph by day of the week for this period.

```{r}
MyYear <- DailyBirths %>%
  filter(year(date) == 1980) %>%
  mutate(week_day = wday(date, label = TRUE))

MyYear %>%
  ggplot(aes(x=date, y = total, color=week_day)) + geom_line()

```

Finally we read in a list of holidays and overlay that information with our plot.

```{r}
Holidays <- read.csv("http://tiny.cc/dcf/US-Holidays.csv", stringsAsFactors = FALSE) %>%
  mutate(date = as.POSIXct(dmy(date))) %>%
  mutate(week_day = wday(date, label = TRUE)) %>%
  filter(year(date) == 1980)

MyYear %>%
  ggplot(aes(x=date, y = total, color=week_day)) +
    geom_line()  +
    geom_vline(data = Holidays, aes(xintercept = as.numeric(date), color = week_day)) +
    geom_text(data = Holidays, mapping=aes(x = date, y = 11000, label = holiday), angle = 45, size = 3, color = "black")
```