---
title: "Hardee's and Carl's Jr locations"
author: "Oliver Thistlethwaite"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

In this project, we will analyze the locations Hardee's and Carl's Jr. First we'll load the required libraries.

```{r, warning = FALSE, message=FALSE}

library(readr)
library(knitr)
library(ggmap)
library(caret)
library(kernlab)
library(dplyr)

```

Now we'll read in the location data for Hardee's and Carl's Jr. These were obtained for free from the [Poi-Factory](http://www.poi-factory.com) website. 

```{r, warning = FALSE, message=FALSE, cache = TRUE}

Restaurants <- read_csv("Hardee's USA.csv", col_names = FALSE) %>% mutate(type = as.factor("Hardees")) %>%
               rbind( read_csv("Carl's Jr_USA_CAN.csv", col_names = FALSE) %>% mutate(type = as.factor("Carls")) )

names(Restaurants) <- c("long", "lat", "name", "address", "type")

```

Next we'll plot the data on a map.

```{r, fig.width=12, fig.height = 12, message = FALSE, warning = FALSE}
map<-get_map(location='united states', zoom=4)

ggmap(map) + geom_point(aes(x = long, y = lat, color = type), data = Restaurants, size = 1, alpha = 0.5) +
  scale_color_manual(values=c("blue", "red"), labels=c("Hardee's", "Carl's Jr"))

```

# Classification

An interesting problem is to see if we can classify whether a restaurant will be Hardee's or Carl's Jr based on its location. We'll use the *Caret* library to do this. First we'll create a training set with 2/3 of the data and use the rest for testing.

```{r, warning = FALSE, message = FALSE}

set.seed(3)

index <- 1:nrow(Restaurants)
testindex <- sample(index, trunc(length(index)/3))
testset <- na.omit(Restaurants[testindex,])
trainset <- na.omit(Restaurants[-testindex,])

```

<br>

## K-Nearest Neighbors

First we'll try k-nearest neighbors. We'll use 10-fold cross validation to find k.

```{r}

fitControl <- trainControl(method = 'cv', number = 10, summaryFunction=defaultSummary)
grid <- expand.grid(k = 1:100)


mod <- train(type ~ long + lat, data = trainset, method = "knn",
             tuneGrid = grid, trControl=fitControl)

pred <- predict(mod, testset)
confusionMatrix(pred, testset$type)
```

So this model did quite well.

<br>

## Support Vector Machine

The obvious model to try is a support vector machine.

```{r, warning = FALSE, message = FALSE}

mod <- train(type ~ long + lat, data = trainset, method = "svmRadial")

pred <- predict(mod, testset)
confusionMatrix(pred, testset$type)
```

There was still 1 restaurant we couldn't get correct!


