---
title: "Knoxville Temperature Prediction"
author: "Oliver Thistlethwaite"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

In this project, we will attempt to predict the temperature in Knoxville, TN using the *ts* and *forecast* libraries.

```{r, warning = FALSE, message=FALSE, include = FALSE}

library(readr)
library(knitr)
library(ggplot2)
library(lubridate)
library(tseries)
library(forecast)
library(dplyr)

```

First we'll read in recent temperature data for Knoxville, TN. This was obtained for free from the [National Centers for Environmental Information](https://www.ncei.noaa.gov/) website. We'll use data is from 1/1/2010 to 5/30/2017. For simplicity we'll just focus on daily maximum temperature from one weather station. We'll also use *lubridate* to turn the DATE column into a genuine date column.

```{r, warning = FALSE, message=FALSE, cache = TRUE}

Temperatures <- read_csv("983224.csv") %>%
  filter(STATION_NAME == "KNOXVILLE MCGHEE TYSON AIRPORT TN US") %>%
  select(DATE, TMAX) %>% mutate(DATE = ymd(DATE)) %>%
  filter(DATE > "2010-1-1", DATE < "2017-5-31")

```

Next we'll plot the last decade of temperature data. The yearly cycle is clearly visible.

```{r}

Temperatures %>% ggplot(aes(x=DATE, y=TMAX)) + geom_line() + labs(y = "daily high temperature")

```


<br>

# TBATS

Now we'll use TBATS to construct our model. This will account for annual seasonality as well as weekly seasonality.

```{r}

TempTS <- msts(Temperatures %>% select(-DATE), seasonal.periods=c(7,365.25), start = 2010)
fit <- tbats(TempTS)
fc <- forecast(fit)
plot(fc)

```

This definitely captured the annual seasonality.

Now we'll print the predictions for the next couple days.

```{r}

Predictions <- as.data.frame(fc$mean)
names(Predictions) <- "TMAX"

Predictions <- Predictions %>%
  mutate(DATE = ymd(20170530) + days(row_number())) %>%
  select(DATE, TMAX)

Predictions %>% head(10) %>% kable()
```

The highs were 81 on 5/31, 81 on 6/1, and 84 on 6/2. So at least these results are semi-reasonable.  
