---
title: "Scraping Nuclear Reactors"
author: "Oliver Thistlethwaite"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this project is to study nuclear reactors in Japan. First we'll read in the required libraries.

```{r, message=FALSE, warnings=FALSE}
library(dplyr)
library(knitr)
library(ggplot2)
library(rvest)
library(lubridate)
library(scales)
```

Now we will read in tables of nuclear reactors from Wikipedia [http://en.wikipedia.org/wiki/List_of_nuclear_reactors](http://en.wikipedia.org/wiki/List_of_nuclear_reactors) via the webscraping library, *rvest*.

```{r}
page <- "http://en.wikipedia.org/wiki/List_of_nuclear_reactors"
table_nodes <- page %>%
  read_html() %>%
  html_nodes("table")
table_list <-
  html_table(table_nodes[1:30], fill = TRUE)

```

After some searching, we find the relevant table is at position 25.

```{r}
table <- table_list[[25]]
table %>% head() %>% kable()
```

Now we fix the column names.

```{r}

names(table) <- c("Name",	"Reactor_No", "Type", "Model", "Status", "Net_Capacity", "Gross_Capacity", "Construction_Start_Date", "Commercial_Operation_Date", "Closure")
table <- table[-1, ] # drop the first row

```

Now we use lubridate to make genuine date columns.

```{r, warning=FALSE}

table <- table %>% mutate(Construction_Start_Date = parse_date_time(Construction_Start_Date, "dmy"),
                          Commercial_Operation_Date = parse_date_time(Commercial_Operation_Date, "dmy"),
                          Closure = parse_date_time(Closure, "dmy"))
```

Now we plot the *construction start date* versus the *net capacity* and *type*.

```{r, warning=FALSE}

table  %>% filter(Type != "") %>% 
  ggplot(aes(x = Construction_Start_Date, y = Net_Capacity)) +
  geom_point(aes(color = Type)) +
  labs(x = "Construction Start Date", y = "Net Capacity")

```


Finally we make an informative graphic that shows how long it took between start of construction and commissioning for each nuclear reactor.

```{r, warning=FALSE}

table$num <- 1:nrow(table)

table  %>% filter(Type != "") %>%
  mutate(Name = paste(Name, Reactor_No)) %>%
  ggplot(aes(y = num)) +
  geom_segment(aes(x = Construction_Start_Date, y = Name, xend = Commercial_Operation_Date, yend = Name)) +
  theme(axis.text.y=element_text(size=5)) +
  labs(x = "Date", y = "Reactor Name")

```