---
title: "The Word on the Street"
author: "Oliver Thistlethwaite"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The purpose of this project is to study which words appear most frequently in street addresses. First we will read in the required libraries.

```{r, message=FALSE, warnings=FALSE}
library(dplyr)
library(knitr)
library(tm)
library(wordcloud2)
```

Now we will read in the street addresses contained in the Medicare database found at  [http://tiny.cc/dcf/CMS_ProvidersSimple.rds](http://tiny.cc/dcf/CMS_ProvidersSimple.rds).

```{r}
Addresses <- readRDS("Medicare.rds") %>% select(address)
docs <- Corpus(VectorSource(Addresses))
```

Next we will clean the text.

```{r}
toSpace <- content_transformer(function (x , pattern ) gsub(pattern, " ", x))
docs <- tm_map(docs, toSpace, "-")
docs <- tm_map(docs, toSpace, "/")
docs <- tm_map(docs, toSpace, "@")
docs <- tm_map(docs, toSpace, "\\|")
docs <- tm_map(docs, removeNumbers)

# Convert the text to lower case
docs <- tm_map(docs, content_transformer(tolower))
# Remove numbers
docs <- tm_map(docs, removeNumbers)
# Remove english common stopwords
docs <- tm_map(docs, removeWords, stopwords("english"))
# Remove punctuations
docs <- tm_map(docs, removePunctuation)
# Eliminate extra white spaces
docs <- tm_map(docs, stripWhitespace)
```

Now we will extract the word frequency information to a dataframe and then print the 20 most frequent words.

```{r}
dtm <- TermDocumentMatrix(docs)
d <- as.data.frame(as.matrix(dtm) )
colnames(d) <- c("freq")
d$word<-rownames(d)
rownames(d) <- 1:nrow(d)
d <- d %>% arrange(desc(freq)) %>% select(word, freq)

d %>% head(20) %>% kable(row.names=TRUE)
```

Finally we will generate a word cloud with the 200 most common words.

```{r}

d <- d %>% head(200)

wordcloud2(d, size = 2)
```




